Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-07-19 paper/paper.tex
Monday WIP, figure generation from sims in a repeatable way

The simulated tectum expresses the same types of receptors and ligands, also
in orthogonal pairs of gradients. In this work, we did not consider the effect
of reverse signalling from tectal receptors to RGC ligands so we only modelled
tectal ligand expression, $L_i$. Although several studies model tectal ligand
expression with exponential functions \citep{koulakov_stochastic_2004}, the
experimental evidence for ligand expression is more ambiguous (Supp. Fig
\textbf{X}). Correspondingly, we leave open the possibility that tectal ligand
expression may be modelled by exponential, linear or logarithmic functions:
\begin{equation}
L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
\end{equation}
or
\begin{equation}
L_i^{\text{lin}}(x,y) = 1.31 + 2.33 x;
\end{equation}
or
\begin{equation}
L_i^{\text{log}}(x,y) = 2.32 + 1.29 \log (2.3(x+0.2))
\end{equation}
(each of which has approximately the same expression values at $x=0$ and $x=1$).

@@

First, we describe the behaviour of a model which has only the gradient-based
chemoaffinity effect and no axon-axon interaction or border effect
(i.e.~$m_2=m_3=0$). This model is run with $n=8$ branches per RGC axon. All
receptor-ligand signals induce repulsive interactions (i.e.~gradient
descent). Receptor and ligand expressions are as shown in
Fig.\,\ref{f:ex}. Fig.\,\ref{f:ch}A gives a colour map of retinal ganglion
cell soma locations, with the colour red encoding $x$ position on the
temporal-nasal axis (with red signifying nasal positions). Green encodes $y$
position on the dorso-ventral axis.
%
Fig.\,\ref{f:ch}A+ shows the expected final location of each RGC axon,
according to experimental observations in a `fishnet' plot, which draws lines
between adjacent axons.
%
Fig.\,\ref{f:ch}B shows the position of
400 branches at $t=500$ (in arbitrary time units). Considering each
coordinate axis separately, each branch interacts differentially with two
opposing gradients. Branches which settle nearest the rostral edge of the
tectum interact more strongly with the gradient $L_0$ than with $L_2$
(Fig\,\ref{f:ex}C). A similar pair of gradients separate branches along the
M-L tectal axis. The branches become positioned on a grid due to the
discretization of the tectal ligand gradient (as the number of branches and
tectal elements tends towards infinity, the arangement tends towards a
continuous map---see Fig.\,S1 \textbf{FIXME}).

@@

Fig.\,\ref{f:ch}C shows the RGC axon centroids; the mean location of $n$
branches per axon. Because all $n$ branches for each RGC have identical
receptor expression and they experience only the chemoaffinity effect, each
branch attains almost the same final position and so the centroid graph
strongly resembles the graph of all the branches. Fig.\,\ref{f:ch}C also shows

>>>>

The simulated tectum expresses the same types of receptors and ligands, also
in orthogonal pairs of gradients. In this work, we did not consider the effect
of reverse signalling from tectal receptors to RGC ligands so we only modelled
tectal ligand expression, $L_i$. Although several studies model tectal ligand
expression with exponential functions \citep{koulakov_stochastic_2004}, the
experimental evidence for ligand expression is more ambiguous (Supp. Fig
\textbf{X}). Correspondingly, we leave open the possibility that tectal ligand
expression may be modelled by exponential, linear or logarithmic functions:
\begin{equation} \label{e:tecexp}
L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
\end{equation}
or
\begin{equation}
L_i^{\text{lin}}(x,y) = 1.31 + 2.33 x;
\end{equation}
or
\begin{equation}
L_i^{\text{log}}(x,y) = 2.32 + 1.29 \log (2.3(x+0.2))
\end{equation}
Each function has approximately the same expression values at $x=0$ and
$x=1$. Opposing functions are obtained by substituting $(1-x)$ for $x$.

@@

\subsubsection*{Wildtype}

First, we describe the behaviour of a model which has only the gradient-based
chemoaffinity effect and no axon-axon interaction or border effect
(i.e.~$m_2=m_3=0$). This model was run with $n=8$ branches per RGC axon. All
receptor-ligand signals induced repulsive interactions (i.e.~gradient
descent). Receptor and ligand expressions were as shown in Fig.\,\ref{f:ex},
noting that tectal ligand expression was of exponential form
(Eq.\,\ref{e:tecexp}). Fig.\,\ref{f:ch}A gives a colour map of retinal
ganglion cell soma locations, with the colour red encoding $x$ position on the
temporal-nasal axis (with red signifying nasal positions). Green encodes $y$
position on the dorso-ventral axis.
%
Fig.\,\ref{f:ch}B shows the expected final location of each RGC axon,
according to experimental observations, in a `fishnet' plot which draws lines
between adjacent axons giving a visual indication of the topology of the
arrangement.
%
Fig.\,\ref{f:ch}C shows the position of
400 branches at $t=500$ (in arbitrary time units). Considering each
coordinate axis separately, each branch interacts differentially with two
opposing gradients. Branches which settle nearest the rostral edge of the
tectum interact more strongly with the gradient $L_0$ than with $L_2$
(Fig\,\ref{f:ex}C). A similar pair of gradients separate branches along the
M-L tectal axis. The branches become positioned on a grid due to the
discretization of the tectal ligand gradient (as the number of branches and
tectal elements tends towards infinity, the arangement tends towards a
continuous map---see Fig.\,S1 \textbf{FIXME}).

@@

Fig.\,\ref{f:ch}D shows a fishnet plot of the RGC axon centroids; the mean
location of $n$ branches per axon. Because all $n$ branches for each RGC have
identical receptor expression and they experience only the chemoaffinity
effect, each branch attains almost the same final position and so the centroid
graph strongly resembles Fig.\,\ref{f:ch}B. Fig.\,\ref{f:ch}D indicates that
the topology of the arrangement resembles that of the experimental result, but
the area over which the branches arrange themselves is limited to about 50\%
of the tectum. It is possible to tailor the tectal ligand expression, whcih
adjusts the relative contributions of the opposing interactions, to tune the
final map. This allows the chemoaffinity model to attain a good match to the
experimental arrangement (Fig.\~S2 \textbf{fixme}). However, the match is
highly dependent on the form of receptor/ligand expression in both retinal and
tectal cells.

\subsubsection*{Tectal ablation etc}
